### 2019-05-30 thesis-template/msc-thesis.tex
update

% Choose language
\newcommand{\setlang}[1]{\selectlanguage{#1}\nonfrenchspacing}
\newcommand*\mean[1]{\bar{#1}}

\begin{document}

@@

Tensorfelder werden meistens in Verbindung mit mechanischen Spanunngsverteilungen in 2D/3D-Gittern gebracht (vgl. Cauchy-Stress Tensor), haben aber auch andere praktische Bedeutungen in der Physik.
Mithilfe von globalen Beleuchtungsmodellen/-techniken wird eine neue Methode entwickelt, um Tensorfelder zu visualisieren, die aber zusätzlich auch verwendet werden kann um die Lichtausbreitung in einer topologischen Szene zu beschreiben. Als Grundlage leiten wir ein einfaches Lichtausbreitungsschema für kartesische Gitter ab, dass die Prinzipien der Ausbreitungsdämpfung und Energieerhaltung beachtet und Licht-verteilung/en für gegebene Lichtquellenrichtung/en und -positionen bis zur Konvergenz approximieren kann. Als unterliegendes Modell werden die Transmissionsprofile innerhalb diesem Gitter als Kristallfiberstruktur angenommen (vgl. Edelsteine: Katzenaugeneffekt beim Tigerauge). Die folgende Aufgabe ist es, anisotrope Fiberstrukturen mit der Orientierung und den Anisotropiemaßen des unterliegenden Tensorfelds zu modellieren. Dafür erstellen wir per Hauptachsentransformation (PCA) ein Eigensystem für jede Zelle des Tensorfelds und modellieren eine zugehörige Ellipsengleichung als (Transfer-/Wichtungs-) Transmissionsprofil. Folglich kann die resultierende Lichtverteilung als 2D Skalarfeld oder Polarplots für Richtungsinformation visualisiert/repräsentiert werden. Wir messen Impulsantworten des Tensorfelds mit Delta-Pulsen an jeder Position und in jede Richtung um eine uniform gesamplete Map der globalen Lichtverteilungen zu erhalten. Diese Map wird als ein globales Energieflussfeld betrachtet. Wir lassen uns von der Technik FTLE aus dem Bereich Particle Tracing der Vector Field Visualization inspirieren, indem wir den Gradient des Flussfelds der resultierenden Lichtverteilungen analysieren und visualisieren. Ähnlich wie FTLE in Vektorfeldern Gräte/Kämme/LCS detektiert, erfasst unsere Methode dieselben Strukturen, anziehende/abstoßende/sattelnde Tensorfeldlinien die häufig Schlüsselstrukturen in der Natur repräsentieren, in Tensorfeldern. Wir führen diese Größe als Global Illumination Gradient ein, der einen FTLE-verwandten Ansatz darstellt, um LCS in Tensorfeldern durch die Analyse des globalen, durch einen Imprint (Gravur in kristallinen Fiberstrukturen) gerichteten Lichtflusses, zu visualisieren. Außerdem führen wir Deformation Ellpsoid Glyphen ein, die zusätzlich zur Darstellung der anisotropen Topologie die Richtung(Zug-/Druck) der Spannung durch Pfeile nach innen/außen darstellen. Das Verfahren wird in einer extensiven Evaluationsphase auf Plausibilität und Vergleichbarkeit geprüft.
\newpage

\chapter*{Abstract}
%\input{abstract}
Tensor Fields, are most commonly associated with stress distributions (cnf. Cauchy-Stress Tensor) in 2D/3D grids, but also have some other meanings in practice in physics. By means of global illumination techniques, we develop a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in a topologically defined scene geometry. As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles and is able to approximate light distributions for given light source position(s) and direction(s) until convergence. The transmission profiles within this grid are considered as crystal fiber structures as an underlying physical model (eg. gemstones: tiger's eye's cat's eye effect). The consequent task is to model anisotropic fibre structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
For this, we derive an eigenbasis by PCA (principal component analysis) for every cell of the tensor field and form an ellipsoid equation as transmission (transfer/weighting) function for the transmitted light profiles. Hence, the resulting light distribution can be visualized via 2D scalar field or polar plots for directional information. Additionally, We measure impulse responses of the tensor field with Delta-Pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions. This map is considered as a global illumination energy flow field.
We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field of the resulting light distributions. Similar to how the FTLE detects ridges/LCS (Lagrangian Coherent Structures) in vector fields (flow fields), our approach captures these structures likewise, attracting/repelling/saddling tensor field lines (TFL) which frequently represent key structures in nature, in tensor fields. We denote this entity a global illumination gradient which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global and light transport  directed by imprint/engraving in crystelline fiber structures. 
We also introduce Deformation Ellipsoid Glyphs which are, additionally to visualizing anisotropic topology, pointing inward/outward for compressive/tensile stresses visualizing orientation of principal mechanical stresses.
We evaluate our approaches for plausibility and comparability within an extensive test campaign.
\newpage

@@

In this work, we wil discuss the advantages of tensor field analysis and develop a new method for tensor field and a coherent light transport visualization. At first, we will shortly state the problem setting objectives and then give a quick overview about the structure of this work.

%Dieses Kapitel gibt einen Überblick über die Arbeit. Gerade der
%Abschnitt zur Motivation soll allgemein verständlich geschrieben
%werden. Die Einleitung sollte auch wichtige Referenzen enthalten.

@@

Tensor Field Visualization is gaining importance as a relevant tool for the analysis of fluid and solid mechanics. Tensors are found most commonly in medical, scientific and engineering applications. A tensor is a $n$-D generalization of a matrix and appears e.g. as Jacobi-Matrix in Vector (Flow) Field Visualization or as stress tensor in solid continuum mechanics. Basically, tensors describe compressions, tensions, rotations and volume changes in both solid and fluid material. While most techniques focus on symmetric tensor field visualization (like Glyphs and Tensor Field Lines), some other recent works adress the problem of asymmetric tensor field visualization. The issue with asymmetric or antisymmetric tensors is that they do not always yield real eigenvalues, which is needed to determine an eigenbasis to set up glyphs, hyperstreamlines or tensor field lines. \cite{pang&zheng} et.al. proposed the concept of dual eigenvectors and \cite{zhang} extended it to pseudo-eigenvectors and introduced eigenvector and eigenvalue manifold to visualize eigenvectors in the complex domain. In this work, we will use the $\mathbf{U}$-matrix obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA principal ellipsoid as suggested by \cite{moler}.\\

\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/DIFFERENT_TYPES_OF_STRESS.png}
  }
  \caption{Different types of stresses}
  \label{Skalierung}
\end{figure}

@@

\section{Objectives}
\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/hyperstreamlines.png}
  }
  \caption{Hyperstreamlines in Tensor Field Visualization}
  \label{Skalierung}
\end{figure}
The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies e.g. clustering, structuring or projection techniques. When it comes to tensor fields, this comes handy in e.g. medical sciences such as in DT-MRI (diffusion tensor-medical resonance imaging) or in simulated (or measured) stress tensor fields from solid and fluid mechanics. The problem setting encomprises the visualization of tensor fields, i.e. mostly 2D/3D-grids consisting of mainly second-order tensors (cnf. Cauchy-Stress Tensor) with the additional possibility to model light transport transport in a geometrical scene. The objective of this work is to develop a simple light propagation scheme following the most basic but crucial physical principles and laws. This propagation scheme can either be used to visualize the light transport paths (separatrices) in a geometry scenery. Or, on the other hand, be used to visualize the anisotropy characteristics of a given tensor field. Therefore the tensor field somehow needs to be pre-processed to generate a tensor-induced footprint map of transmission profiles. So however, the tensors need to be decomposed to obtain a kind of transmission profile. Another aim is to implement an approach borrowed from vector-field visualization. We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges/LCS (Lagrangian Coherent Structures). These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of differing stress distributions (w.r.t. direction - measuring anisotropy/location - measuring ). The map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions. The real part of the eigenvalues can be exploited for sign change (contractions/tensions) which leads to the concept of deformation glyphs w. attached arrows pointing inward/outward for the two types of normal stresses. Practically, in its eigenbasis there are no shear stresses existent for the symmetric stress tensor in equilibrium, just the principal normal stresses. Thus, we can even define the full set of stresses with the deformation glyphs which are derived from the singular value decomposition and signed with the real part of the eigenvalues (in corresponding absolute order).

@@

\section{Structure of this work}

@@

In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and will capture Tensors, PCA and Related Work in that sense.\\

@@

\section{Tensors}
\label{sec:tensors}
Now what is a Tensor? A tensor is a $o$-dimensional generalization of a matrix as follows:
\begin{center}
 \begin{tabular}{c|c|c|c|c|c|c}
\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
\hline 
\textbf{shape} & scalar & vector & matrix & "3D-matrix" & ... & oD-matrix\\ 
\end{tabular}\vskip 5pt
Table 1: Tensor Shapes
 \end{center}

In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ (the dimensin of the row and column vectors), which is equal to the rank of the tensor for full rank (or in other words the number of elements n for a $nxn$-matrix). Tensors follow certain transformation rules which are defined for covariant/contravariant tensors. They are most commonly obtained and defined from physical transformation equations. Hence, any $o$-D number array could be a tensor, but the definition holds only if the transformation rules apply to these $o$-D number arrays (in a sense of physical interpretation), which is mostly true for component-indexed $o$-D number entities following matrix multiplication and transformation rules found in physics and math. In fact, tensors themselves are interpreted as a transformation which reflects an incoming direction vector (measuring e.g. the stress in that direction) to an outgoing stress vector.
 To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities aren't simple numbers or vectors. They occur in form of a tensor fields, and thus as entities or quantities in physics. That is, at each point in space (the grid) there is a whole distribution of e.g. stresses (elastic and viscous) which needs to be characterized and represented by a single tensor.
 \\
 \emph{Cauchy stress tensor}
 \begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/stress_tensor.png}
  }
  \caption{Cauchy-Stress Tensor}
  \label{cauchy}
\end{figure}

The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig. \ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$ arranged in row-major order. It appears in similar form as viscous stress tensor in fluid mechanics:
\begin{align*}
\boldsymbol{\sigma_{ij}}
 = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
\boldsymbol{T}^{(\mathbf{e}_2)} \\
\boldsymbol{T}^{(\mathbf{e}_3)} \\
\end{matrix}}\right] =
\left[{\begin{matrix}
\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
\end{matrix}}\right]  = \left[{\begin{matrix}
\sigma _x & \tau _{xy} & \tau _{xz} \\
\tau _{yx} & \sigma _y & \tau _{yz} \\
\tau _{zx} & \tau _{zy} & \sigma _z \\
\end{matrix}}\right] =
\begin{bmatrix}
\sigma _{11} & \sigma _{12} & \sigma _{13} \\
\sigma _{21} & \sigma _{22} & \sigma _{23} \\
\sigma _{31} & \sigma _{32} & \sigma _{33} 
\end{bmatrix}
\end{align*}
These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in direction ${x,y,z}$ (9 individual numbers). Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space. The stress tensor itself can be interpreted as a transformation which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$. The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent. It is possible to find 3 principal stress directions in 3D through PCA (see sect. \ref{sec:pca} below), for which there are no shear stresses existent in equilibrium. These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour. The resulting system of principal directions and absolute stresses is often called an "eigensystem" or "eigenframe". So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig. \ref{HAT} in sect. \ref{sec:pca}.

\section{Principal Component Analysis}
\label{sec:pca}
Note that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following. For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by e.g. ellipsoid glyphs.

\begin{figure}[h]
%\centering
  \begin{minipage}{0.5\textwidth}
    \includegraphics[width=\textwidth]{img/HAT.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \includegraphics[width=\textwidth]{img/glyphs.png}
    \label{b)}
  \end{minipage}
\caption{a) 2D Principal Component Analysis, b) 3D Ellipsoid Glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
\label{gradient}
\end{figure}

Principal Component Analysis is an algorithm which can capture the principal components (directions) in stochastic data. But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling. Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices. For (skew-)symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cnf. \cite{kieburg}). Note that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases. Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling ($\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of matrix as pointed out by \cite{moler}. In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues (\cite{mroz}) ordered decreasingly by absolute value (corresponding to singular value order). The sign corresponds to tensile ($+$) and compressive ($-$) stress, which would allow us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses. We do this decomposition analysis once for each matrix and then store the precomputed results in a grid. The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step. In this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition. This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor. We will then use global-illumination techniques on top to generate a flow map, which permits us to compute an FTLE-like field on the tensor fields.

\section{GT Algorithms / Approaches}
\label{sec:gt}
\subsection{Glyphs and Tensorfieldlines}
\begin{figure}[h]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/tensorfieldlines.png}
  }
  \caption{white: Tensorfieldlines; green: Ellipsoid Glyphs}
  \label{scheme}
\end{figure}\vskip 3pt
Glyphs were derived as a basis for the weighting profiles of the tensors in the following propagation scheme. They were chosen because it is an easy-to-use (understand) concept which can also be interpreted nicely. In fact, these are used for the principal component analysis of direction vectors for the Tensor Field Line extrapolation (integration) in a subsequent step.
Tensor Field Lines (the type of tensor stream line - hyperstreamline analog that does not introduce artificial intertia) were implemented for Ground Truth (GT) data generation purposes. For example, when we design an FTLE-like field computation approach, we will need to know when tensor field lines diverge/converge.

\subsection{Test Data (Generation)}
We used python's numpy library to generate pre-defined test cases and caught one real data example from \url{TensorVis.org}, which needed to be pre-processed through slicing and subsequent cropping to 2D samples (2x2 matrices), which can be grouped into symmetric and asymmetric ones:\\

>>>>

% Choose language
\newcommand{\setlang}[1]{\selectlanguage{#1}\nonfrenchspacing}
\newcommand*\mean[1]{\bar{#1}}\begin{document}

@@

Tensorfelder werden meistens in Verbindung mit mechanischen Spanunngsverteilungen in 2D/3D-Gittern gebracht (vgl. Cauchy-Stress Tensor), haben aber auch andere praktische Bedeutungen in der Physik.
Mithilfe von globalen Beleuchtungsmodellen/-techniken wird eine neue Methode entwickelt, um Tensorfelder zu visualisieren, die aber zusätzlich auch verwendet werden kann um die Lichtausbreitung in einer topologischen Szene zu beschreiben. Als Grundlage leiten wir ein einfaches Lichtausbreitungsschema für kartesische Gitter ab, dass die Prinzipien der Ausbreitungsdämpfung und Energieerhaltung beachtet und Licht-verteilung/en für gegebene Lichtquellenrichtung/en und -positionen bis zur Konvergenz approximieren kann. Als unterliegendes Modell werden die Transmissionsprofile innerhalb diesem Gitter als Kristallfiberstruktur angenommen (vgl. Edelsteine: Katzenaugeneffekt beim Tigerauge). Die folgende Aufgabe ist es, anisotrope Fiberstrukturen mit der Orientierung und den Anisotropiemaßen des unterliegenden Tensorfelds zu modellieren. Dafür erstellen wir per Hauptachsentransformation (PCA) ein Eigensystem für jede Zelle des Tensorfelds und modellieren eine zugehörige Ellipsengleichung als (Transfer-/Wichtungs-) Transmissionsprofil. Folglich kann die resultierende Lichtverteilung als 2D Skalarfeld oder Polarplots für Richtungsinformation visualisiert/repräsentiert werden. Wir messen Impulsantworten des Tensorfelds mit Delta-Pulsen an jeder Position und in jede Richtung um eine uniform gesamplete Map der globalen Lichtverteilungen zu erhalten. Diese Map wird als ein globales Energieflussfeld betrachtet. Wir lassen uns von der Technik FTLE aus dem Bereich Particle Tracing der Vector Field Visualization inspirieren, indem wir den Gradient des Flussfelds der resultierenden Lichtverteilungen analysieren und visualisieren. Ähnlich wie FTLE in Vektorfeldern Gräte/Kämme/LCS detektiert, erfasst unsere Methode dieselben Strukturen, anziehende/abstoßende/sattelnde Tensorfeldlinien die häufig Schlüsselstrukturen in der Natur repräsentieren, in Tensorfeldern. Wir führen diese Größe als Global Illumination Gradient ein, der einen FTLE-verwandten Ansatz darstellt, um LCS in Tensorfeldern durch die Analyse des globalen, durch einen Imprint (Gravur in kristallinen Fiberstrukturen) gerichteten Lichtflusses, zu visualisieren.
\newpage

\chapter*{Abstract}
%\input{abstract}
Tensor fields, are most commonly associated with stress distributions (cf. Cauchy-stress tensor) in 2D/3D grids, but also have some other meanings in practice in physics. By means of global illumination techniques, we motivate a new method to visualize tensor fields, which is also capable of visualizing the light transport (propagation) in a topologically defined scene geometry. As a basis, we derive a simple light propagation scheme for Cartesian grids, which satisfies propagation attenuation and energy-conservation principles and is able to approximate light distributions for given light source position(s) and direction(s) until convergence. The transmission profiles within this grid are considered as crystal fiber structures as an underlying physical model (eg. gemstones: tiger's eye's cat's eye effect). The consequent task is to model anisotropic fibre structures with the orientation and anisotropy measures of the underlying tensor field in the grid.
For this, we derive an eigenbasis by PCA (principal component analysis) for every cell of the tensor field and form an ellipsoid equation as transmission (transfer/weighting) function for the transmitted light profiles. Hence, the resulting light distribution can be visualized via 2D scalar field or polar plots for directional information. Additionally, We measure impulse responses of the tensor field with Dirac-pulses as light sources at any (sampled) position and in any direction to generate a uniform sampled map of the global illumination distributions. This map is considered as a global illumination energy flow field.
We gain inspiration by vector field particle tracing's FTLE approach for analyzing the gradient of the flow field of the resulting light distributions. Similar to how the FTLE detects ridges/LCS (Lagrangian coherent structures) in vector fields (flow fields), our approach captures these structures likewise, attracting/repelling/saddling tensor field lines (TFL) which frequently represent key structures in nature, in tensor fields. We denote this entity a global illumination gradient which states an FTLE-related approach for visualizing LCS in tensor fields through analysis of the global and light transport  directed by imprint/engraving in crystelline fiber structures modeled by the tensor field's ellipsoid glyhps. 
We also introduce Deformation Ellipsoid glyphs which are, additionally to visualizing anisotropic topology, pointing inward/outward for compressive/tensile stresses visualizing orientation of principal mechanical stresses.
We evaluate our approaches for plausibility and comparability within an extensive test campaign.
\newpage

@@

In this work, we will discuss the advantages of tensor field analysis and motivate a new method for tensor field through a coherent light transport visualization. At first, we will shortly state the problem setting objectives and then give a quick overview about the structure of this work.

\paragraph{Contributions of this Work}
\begin{itemize}
	\item
\end{itemize}
%Dieses Kapitel gibt einen Überblick über die Arbeit. Gerade der
%Abschnitt zur Motivation soll allgemein verständlich geschrieben
%werden. Die Einleitung sollte auch wichtige Referenzen enthalten.

@@

Tensor Field Visualization is gaining importance as a relevant tool for the analysis of fluid and solid mechanics. Tensors are found most commonly in medical, scientific and engineering applications. A tensor is a $n$-D generalization of a matrix and appears, e.g., as Jacobian-Matrix in vector (Flow) field visualization or as stress tensor in solid continuum mechanics. Basically, tensors describe compressions, tensions, rotations and volume changes in both solid and fluid material. While most techniques focus on symmetric tensor field visualization (like glyphs and tensor field lines), some other recent works adress the problem of asymmetric tensor field visualization. The issue with asymmetric or antisymmetric tensors is that they do not always yield real eigenvalues, which is needed to determine an eigenbasis to set up glyphs, hyperstreamlines or tensor field lines. \cite{pang&zheng} et al. proposed the concept of dual eigenvectors and \cite{zhang} extended it to pseudo-eigenvectors and introduced eigenvector and eigenvalue manifold to visualize eigenvectors in the complex domain. In this work, we will use the $\mathbf{U}$-matrix obtained from singular value decomposition (SVD) to obtain the half-axes of the PCA principal ellipsoid as suggested by \cite{moler}.\\

\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/DIFFERENT_TYPES_OF_STRESS.png}
  }
  \caption{Different types of stresses}
  \label{Skalierung}
\end{figure}

@@

\section{Objectives}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/hyperstreamlines.png}
  }
  \caption{Hyperstreamlines in Tensor Field Visualization}
  \label{Skalierung}
\end{figure}
The main purpose of visualization in general is to reduce or encode some set of numbers in some readable, explorable and more intuitive representation. This implies, e.g., clustering, structuring or projection techniques. When it comes to tensor fields, this comes handy in, e.g., medical sciences such as in DT-MRI (diffusion tensor-medical resonance imaging) or in simulated (or measured) stress tensor fields from solid and fluid mechanics. The problem setting encomprises the visualization of tensor fields, i.e., mostly 2D/3D-grids consisting of mainly second-order tensors (cf. Cauchy-Stress Tensor) with the additional possibility to model light transport transport in a geometrical scene. The objective of this work is to design a simple light propagation scheme following the most basic but crucial physical principles and laws. This propagation scheme can either be used to visualize the light transport paths (separatrices) in a geometry scenery. Or, on the other hand, be used to visualize the anisotropy characteristics of a given tensor field. Therefore the tensor field somehow needs to be pre-processed to generate a tensor-induced footprint map of transmission profiles. So however, the tensors need to be decomposed to obtain a kind of transmission profile. Another aim is to implement an approach borrowed from vector-field visualization. We construct an FTLE-related approach in tensor field visualization (both symmetric and asymmetric) for detecting ridges/LCS (Lagrangian coherent structures). These structures can be classified (identified) as repelling, attracting and saddling (forming a saddle) local features, which frequently represent key structures in nature. LCS usually separate regions of different flow behavior, but in our case they separate regions/domains of differing stress distributions (w.r.t. direction - measuring anisotropy/location - measuring ). The map responds most where tensor field lines / hyperstreamlines converge or diverge, whereas one is the same as the other because of bidirectional stress directions. The real part of the eigenvalues can be exploited for sign change (contractions/tensions) which leads to the concept of deformation glyphs w. attached arrows pointing inward/outward for the two types of normal stresses. Practically, in its eigenbasis there are no shear stresses existent for the symmetric stress tensor in equilibrium, just the principal normal stresses. Thus, we can even define the full set of stresses with the deformation glyphs which are derived from the singular value decomposition and signed with the real part of the eigenvalues (in corresponding absolute order).

@@

\section{Structure of this Work}

@@

In this chapter, we will capture the prerequisites necessary to understand the contributions of this work and will capture Tensors, PCA and Related Work in that sense.

@@

\section{Tensor Fields}
\label{sec:tensors}
Now what is a Tensor? A tensor is a $o$-dimensional generalization of a matrix as illustrated in Table \ref{tensor}.
\begin{table}[!b]
\centering
\caption{Tensor Shapes}
 \begin{tabular}{c|c|c|c|c|c|c}
\textbf{order} & 0 & 1 & 2 & 3 & ... & o \\ 
\hline 
\textbf{shape} & scalar & vector & matrix & "3D-matrix" & ... & oD-matrix\\ 
\end{tabular}
\label{tensor}
 \end{table} 
In addition to that, a tensor has as many indices as its order $o$ and their run length is as long as its embedded dimension $n$ (the dimensin of the row and column vectors), which is equal to the rank of the tensor for full rank (or in other words the number of elements n for a $nxn$-matrix). Tensors follow certain transformation rules which are defined for covariant/contravariant tensors. They are most commonly obtained and defined from physical transformation equations. Hence, any $o$-D number array could be a tensor, but the definition holds only if the transformation rules apply to these $o$-D number arrays (in a sense of physical interpretation), which is mostly true for component-indexed $o$-D number entities following matrix multiplication and transformation rules found in physics and math. In fact, tensors themselves are interpreted as a transformation which reflects an incoming direction vector (measuring, e.g., the stress in that direction) to an outgoing stress vector.
 To be a bit more concrete, tensors represent stress distributions in solids and fluids describing strain and flow features as these entities aren't simple numbers or vectors. They occur in form of a tensor fields, and thus as entities or quantities in physics. That is, at each point in space (the grid) there is a whole distribution of, e.g., stresses (elastic and viscous) which needs to be characterized and represented by a single tensor.
 \\
 \paragraph{stress tensor}
 \begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/stress_tensor.png}
  }
  \caption{Cauchy-Stress Tensor}
  \label{cauchy}
\end{figure}

The Cauchy stress tensor, which is propably the classical example of a tensor, is depicted in Fig. \ref{cauchy} and consists of 3 stress vectors $\mathbf{T(e_i)}$ arranged in row-major order. It appears in similar form as viscous stress tensor in fluid mechanics:
\begin{align*}
\boldsymbol{\sigma}
 = \left[{\begin{matrix} \mathbf{T}^{(\mathbf{e}_1)} \\
\boldsymbol{T}^{(\mathbf{e}_2)} \\
\boldsymbol{T}^{(\mathbf{e}_3)} \\
\end{matrix}}\right] =
\left[{\begin{matrix}
\sigma _{xx} & \sigma _{xy} & \sigma _{xz} \\
\sigma _{yx} & \sigma _{yy} & \sigma _{yz} \\
\sigma _{zx} & \sigma _{zy} & \sigma _{zz} \\
\end{matrix}}\right]  = \left[{\begin{matrix}
\sigma _x & \tau _{xy} & \tau _{xz} \\
\tau _{yx} & \sigma _y & \tau _{yz} \\
\tau _{zx} & \tau _{zy} & \sigma _z \\
\end{matrix}}\right] =
\begin{bmatrix}
\sigma _{11} & \sigma _{12} & \sigma _{13} \\
\sigma _{21} & \sigma _{22} & \sigma _{23} \\
\sigma _{31} & \sigma _{32} & \sigma _{33} 
\end{bmatrix}
\end{align*}.
These 3 stress vectors represent the orientation and magnitude of total resulting stress at plane ${x,y,z}$ in direction ${x,y,z}$ (9 individual numbers). Thus, the stress tensor poses a full valid represenation of the stress distribution at any point in space. The stress tensor itself can be interpreted as a transformation which maps an incoming normal direction vector $\mathbf{n}$ a resulting stress vector $\mathbf{T^{(n)}} = \mathbf{n}\cdot\boldsymbol{\sigma}$. The stresses in normal directions (diagonal elements) are directly related to the pressure, which is always isotropic and hence non-directionally dependent. It is possible to find 3 principal stress directions in 3D through PCA (see sect. \ref{sec:pca} below), for which there are no shear stresses existent in equilibrium. These form the principal directions of deformation at a certain location and are sufficient to describe the resulting transformation/deformation behaviour. The resulting system of principal directions and absolute stresses is often called an "eigensystem" or "eigenframe". So to say, an eigensystem, consistent of the principal directions/axes, spans a principal ellipsoid depicted in Fig. \ref{HAT} in sect. \ref{sec:pca}.

\section{Principal Component Analysis}
\label{sec:pca}
Note that we denote a matrix with a second order tensor interchangeably for the sake of simplicity in the following. For any state of stress in equilibrium it is possible to find $n$ (dimensionality) independent orthogonal directions with no shear stresses existent. Along these directions the principal stresses, which are normal (tensile/compressive) stresses, are exerted. The anisotropy can be represented by, e.g., ellipsoid glyphs.

\begin{figure}[!t]
%\centering
  \begin{minipage}{0.5\textwidth}
    \includegraphics[width=\textwidth]{img/HAT.png}
    \label{a)}
  \end{minipage}
  \begin{minipage}{0.5\textwidth}
    \includegraphics[width=\textwidth]{img/glyphs.png}
    \label{b)}
  \end{minipage}
\caption{a) 2D Principal component analysis, b) 3D ellipsoid glyphs w. anisotropy coefficient (l:linear, p:planar, s:spherical)}
\label{gradient}
\end{figure}

Principal Component Analysis is an algorithm which can capture the principal components (directions) in stochastic data. But it works similarly on transformations to capture the principal directions from transformations like rotation and scaling. Now, principal component analysis can be done by eigenvalue or singular vector decomposition, whereas the former yields complex eigenvalues for asymmetric matrices. For (skew-)symmetric (normal) matrices, they are connected through the following relation: $s_i = \sqrt{|\lambda_i|}$ (cf. \cite{kieburg}). Note that the eigenvectors match the singular vectors only for symmetric matrices, which does not mean that they do not form the half-axes of the principal component ellipsoid in other cases. Mathematically speaking, a decomposition is somehow decomposing a transformation (which follows certain transformation rules) into parts of subsequent transformations consistent of rotation and scaling ($\mathbf{A} = \mathbf{R}\mathbf{S}\mathbf{R}^*$). We choose to use singular value decomposition ($\mathbf{A} = \mathbf{U} \mathbf{\Sigma} \mathbf{V}^*$) to be able to set up real eigenvalues for any kind of matrix (singular values are always real and positive). Also, the singular vectors of the $\mathbf{U}$-matrix represent the half-axes of the principal ellipsoid (eigenframe) shown in Fig. \ref{HAT} for any kind of matrix as pointed out by \cite{moler}. In addition, we also calculate the eigenvalues to exploit the sign of the real part of the eigenvalues \cite{mroz} ordered decreasingly by absolute value (corresponding to singular value order). The sign corresponds to tensile ($+$) and compressive ($-$) stress, which would allow us to visualize deformation glyphs with arrows pointing inwards for compressive and outwards for tensile stresses. We do this decomposition analysis once for each matrix and then store the precomputed results in a grid. The principal ellipsoid is also used as a transmission profile for the propagation scheme in symbolic form in a subequent step. In this work, we will use the singular vectors yielded from singular value decomposition to avoid complex values yielded from eigenvalue decomposition. This will allow us to set up glyphs representing the major/minor axis of the underlying transformation for every kind of (asymmetric/symmetric) tensor. We will then use global-illumination techniques on top to generate a flow map, which permits us to compute an FTLE-like field on the tensor fields.

\section{GT Algorithms / Approaches}
\label{sec:gt}
\paragraph{Glyphs and tensor field lines}
\begin{figure}[!t]
  \centering
 {
    \includegraphics[width=0.6\textwidth]
    {img/tensorfieldlines.png}
  }
  \caption{white: tensorfieldlines; green: ellipsoid glyphs}
  \label{scheme}
\end{figure}\vskip 3pt
Glyphs were derived as a basis for the weighting profiles of the tensors in the following propagation scheme. They were chosen because it is an easy-to-use (understand) concept which can also be interpreted nicely. In fact, these are used for the principal component analysis of direction vectors for the Tensor Field Line extrapolation (integration) in a subsequent step.
Tensor Field Lines (the type of tensor stream line - hyperstreamline analog that does not introduce artificial intertia) were implemented for Ground Truth (GT) data generation purposes. For example, when we design an FTLE-like field computation approach, we will need to know when tensor field lines diverge/converge.

\paragraph{Test Data (Generation)}
We used python's numpy library to generate pre-defined test cases and caught one real data example from \url{TensorVis.org}, which needed to be pre-processed through slicing and subsequent cropping to 2D samples (2x2 matrices), which can be grouped into symmetric and asymmetric ones:\\
